On Optimality of Finite Element Integration

### 2015-12-22 TOMS_Paper/assembly.tex
Improving the proof...

\begin{enumerate}
\item 
\begin{enumerate}
\item Let $T_m$ represent the set of tables resulting from applying pre-evaluation to a monomial $m$. Consider two monomials $m_1, m_2 \in Y$ and the respective sets of pre-evaluated tables, $T_{m_{1}}$ and $T_{m_{2}}$. If $T_{m_{1}} \cap T_{m_{2}} \neq \emptyset$, at least one table is assignable to the same temporary. $\Gamma$, therefore, may skip a potential global optimum, since $X$ and $Y$ are constructed reasoning -- through $\theta$ -- only on monomials in ``isolation''. This case is neglected by assumption. The assumption is reasonable because such a scenario is purely pathological and, with extreme probability, of insignificant importance.
\item Let $m_1 \in X$ and $m_2 \in Y$ be two monomials that share some generic multilinear symbols. If $m_1$ were carelessly pre-evaluated, there may be a potential gain in sharing elimination that is lost, potentially leading to a local -- instead of a global -- optimum. This situation is prevented by construction, because $\Gamma$ analyzes all possible bipartitions searching for a global optimum that also preserves the memory constraints. Recall that since the number of monomials is in practice very small, this pass can rapidly be accomplished. Finally, note how this problem can be seen as an instance of the well-known Knapsack problem.
\end{enumerate}
\item Consider Algorithm~\ref{algo:sharing-elimination}. Based on loop linearity and normal form of input expressions, we observe that there are only two ways of scheduling the multilinear operands in $P \in \mathbb{P}$: through (i) generalized code motion or (ii) factorization of multilinear symbols, respectively leading to $|P|$ and $|S_P|$ multiplications. An upper bound for $P$ is, therefore, $ub_P = 2{min(}|P|,\ |S_P|{)} - 1$. Thus, when generalized code motion in step (3) is applied, we do know that we are not pruning any potentially optimal solution from the search space. In fact, we are just moving towards the lower bound, given by $lb_P = min(0, ub_P)$, with $0$ reached only if the symbols in $P$ will be grouped and hoisted at step (4).

Step (4) factorizes multilinear symbols and temporaries produced at step (3) in optimal order; that is, an order that leads to a quasi-optimal synthesis. To demonstrate this, we consider the sharing graph rewrite rules.
\begin{enumerate}[label=(\Alph*)]
\item Rule (A) has no side-effects since $\lbrace T^0, ..., T^n \rbrace$ can only be grouped through the factorization of $S$.
\item By contradiction, we can prove that applying rule (B) is always a step forward towards the optimal solution, assuming (as stated in Algorithm~\ref{algo:sharing-elimination}) no hit is possible for rule (A). Assume then that (B) prunes the path towards quasi-optimality by choosing a sub-optimal factorization. This means that there exists ...
\item Rule (C) ...
\end{enumerate}
Finally, we note that step (5) is irrelevant for quasi-optimality.

>>>>

\begin{enumerate}
\item 
\begin{enumerate}
\item Let $T_m$ represent the set of tables resulting from applying pre-evaluation to a monomial $m$. Consider two monomials $m_1, m_2 \in Y$ and the respective sets of pre-evaluated tables, $T_{m_{1}}$ and $T_{m_{2}}$. If $T_{m_{1}} \cap T_{m_{2}} \neq \emptyset$, at least one table is assignable to the same temporary. $\Gamma$, therefore, may skip a potential global optimum, since $X$ and $Y$ are constructed reasoning -- through $\theta$ -- only on monomials in ``isolation''. This case is neglected by assumption. The assumption is reasonable because such a scenario is purely pathological and, with extreme probability, of insignificant importance.
\item Let $m_1 \in X$ and $m_2 \in Y$ be two monomials that share some generic multilinear symbols. If $m_1$ were carelessly pre-evaluated, there may be a potential gain in sharing elimination that is lost, potentially leading to a local -- instead of a global -- optimum. This situation is prevented by construction, because $\Gamma$ analyzes all possible bipartitions searching for a global optimum that also preserves the memory constraints. Recall that since the number of monomials is in practice very small, this pass can rapidly be accomplished. Finally, note how this problem can be seen as an instance of the well-known Knapsack problem.
\end{enumerate}
\item Consider Algorithm~\ref{algo:sharing-elimination}. Because of loop linearity and normal form of expressions, there are only two ways of scheduling the multilinear operands in $P \in \mathbb{P}$: through (i) generalized code motion or (ii) factorization of multilinear symbols. These two strategies lead, respectively, to performing $|P|$ and $|S_P|$ products at every iteration. Since (i) is side effects free (i.e., neither expansion nor factorization are needed) and is applied if and only if $|P| < |S_P|$, it is clear that step (3) cannot prune any optimal solution from the search space; in fact, it just moves towards quasi-optimality.

Step (4) factorizes symbols following an ordering that brings to quasi-optimality. The order is dictated by the sharing graph and its rewrite rules. We first recall that rewrite rules are tried following the order in which they were presented in Algorithm~\ref{algo:sharing-elimination}; for example, we do not attempt rule (B) until we have handled all possible hits for rule (A).
\begin{enumerate}[label=(\Alph*)]
\item Rule (A) is side effects free since $\lbrace T^0, ..., T^n \rbrace$ can be grouped only through the factorization of $S$.
\item We prove that rule (B) makes the algorithm progress towards the optimal solution. We first observe that the choice of $S^{nt}$ is arbitrary. Since $S^{nt}$ is no terminal, there must exist at least one edge between $S^{nt}$ and a generic vertex $V$. We then consider the alternative factorization, namely $S^{nt}(...S + ... + ...V)$. If $V$ were a terminal, we would simply be back to where we started, because rule (B) would be applicable by setting $V = T$ and $S = S^{nt}$. If, on the other hand, $V$ were a non terminal, we would have $S$ appearing in two different sub-expressions, which implies the presence of spatial sharing.
\item In order to have $\mathbb{T} = \emptyset$, which would trigger rule (C), it must be the case that all non terminals have the same number of incoming edges.  
\end{enumerate}
Finally, we note that step (5) is irrelevant for quasi-optimality.
